When historical observations, models, and experiments agree: Interspecific competition is weak in a sagebrush steppe

### 2016-10-02 removals/manuscript/removal_ms.tex
adding more revisions to removal manuscript on andy edits branch.  Adding more specific references for Pseudoroegeneria competition, spatial segregation and hydraulic lift

% I could be wrong but I think the left-hand side of the equation needs to be 
% indexed to {t + 1} not t.

@@

%% Wait, how are recruit positions within the plot distributed?

@@

% There’s a bit of a leap here from statistical significance to predictions
% I think this is intuitively correct and probably right most of the time, but 
% I wonder whether a main effect of removal treatment necessarily implies 
% that predictions don’t match observations. (I guess it probably does).  
% However, in the case that treatment effects are not significant, I think there
% could be cases where predictions still technically differ from observations. For % instance, do predictions technically differ from observations if they predict the % mean, but underestimate the variance? Something to consider.

@@

% What about other species’ recruitment results?

@@

%
% Peter the dashed and dotted lines in this figure don’t show up as distinct to me 
% (baseline vs. removal predicted).  I can’t tell which is which
%

@@

% another hypotheses about PSSP
% I predict you will cut this! :)

Another explanation for the under-prediction of \textit{P. spicata} competitive release may come from consideration of the scale at which plots are monitored in this experiment. In our models the competition experienced by a focal plant is weighted by distance of competitors from the focal plant. However because our plots are 1 m across we have no information about the effects of competition at scales greater than on 1 m.  Given that \textit{Artemisia} roots can spread laterally beyond their canopy’s extent (Kleinhesselink personal observation), it is possible that even \textit{P. spicata} plants in plots without \textit{Artemisia} canopies, a situation in theory identical to our removals, are experiencing some competition below ground from \textit{Artemisia} outside the plots. When we conducted removals we removed plants within 50 cm of the removal plots. These “outside the bounds” removals could give an added boost to \textit{P. spicata} performance that would not be captured by our historical data.  Additionally, there is growing evidence that plant roots are territorial, they recognize self and non-self roots and respond differently to the roots of different species \citep{Krinnetz and Caldwell}. Moreover, roots of many species appear to grow towards interspecific competition and avoid intraspecific neighbors—resulting in patterns of intraspecific root aggregation \citep{bertelheimer_aggregative_2006}.  If \textit{Artemisia} roots behave in this way, it could mean that our removal plots contain much lower densities of \textit{Artemisia} roots belowground than control plots without \textit{Artemisia} canopies aboveground. If this is the case, we could find that as \textit{Artemisia} roots from plants outside our plots re-invade the territory within the plots, the growth of \textit{P. spicata} will gradually be reduced and come to match that in control plots without \textit{Artemisia}.

The small competitive release of many species in our system implies that each of the four dominant species we studied are limited by quite different factors. This low niche overlap could reflect strong resource partitioning (Tilman 1982), perhaps combined with variation in the times and locations at which species are taking up those resources. Species-specific responses to spatial and temporal variation in non-resource environmental variables can also weaken interspecific interactions \citep{chesson_mechanisms_2000}, though in a previous study we failed to find evidence of a temporal storage effect in this community \citep{adler_weak_2009}. Interactions with above- or belowground enemies can regulate populations and, given enough species-specificity in the effects of these enemies, decouple interspecific dynamics \citep{hersh_evaluating_2011,janzen_herbivores_1970,connell_role_1971}.  More likely, plants may alter the environment, and even resource availability, in ways that facilitate their neighbors. For example, shade cast by \textit{Artemisia} canopies may reduce evapotranspiration (e.g. \citealt{Barbier2008}). \textit{Artemisia} removal might offest this facilitative effect by reducing resource uptake, resulting in little net change in water balance following removal.  Unfortunately, we did not monitor soil moisture at the fine spatial and temporal resolution necessary to test this hypothesis.

@@

% added this nutnet tie-in that I think is cool; it could go up after the Yenni 
% reference above.

>>>>

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% I could be wrong but I think the left-hand side of the equation above
% needs to be  indexed to {t + 1} not t.  
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Wait, how are recruit positions within the plot distributed? 
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% There’s a bit of a leap here from statistical significance to predictions
% I think this is intuitively correct and probably right most of the time, but 
% I wonder whether a main effect of removal treatment necessarily implies 
% that predictions don’t match observations. (I guess it probably does).  
% However, in the case that treatment effects are not significant, I think there
% could be cases where predictions still technically differ from observations. 
% For % instance, do predictions technically differ from observations if they predict the 
% mean, but underestimate the variance? Something to consider.
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% What about other species’ recruitment results? 
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Peter the dashed and dotted lines in this figure don’t show up as distinct to me 
% (baseline vs. removal predicted).  I can’t tell which is which
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Here's my   hypotheses about PSSP. I predict you will cut this! :)   
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

Another explanation for the under-prediction of \textit{P. spicata} competitive release may come from consideration of the scale at which plots are monitored in this experiment. In our models, the competition experienced by a focal plant is weighted by distance of competitors from the focal plant. However we have no information about the effects of competition at scales greater than on 1 m.  Given that \textit{Artemisia} roots can spread laterally beyond their canopy’s extent (Kleinhesselink personal observation), it is possible that even \textit{P. spicata} plants in plots without \textit{Artemisia} canopies, a situation in theory identical to our removals, are experiencing some competition with \textit{Artemisia} whose roots extend into the plots from outside. When we conducted removals we removed plants within 50 cm of the removal plots. These marginal removals could give an added boost to \textit{P. spicata} performance that would not be captured by the crowding covariate in our model. Additionally, there is evidence from many species, including desert shrubs that plant roots from individuals are often spatially segregated \citep{schenk_spatial_1999, brisson_effect_1994}. If this is the case for \textit{Artemisia}, then the roots within a plot may belong to only one or two \textit{Artemisia} plants. If we remove those individuals, it seems possible that over time roots from other individual \textit{Artemisia} plants outside our plots will invade the removal plots over time. A gradually lessening of the removal effect \textit{P. spicata} in the removal plots over time would be support this hypothesis.

The small competitive release of many species in our system implies that each of the four dominant species we studied are limited by quite different factors. This low niche overlap could reflect strong resource partitioning (Tilman 1982), perhaps combined with variation in the times and locations at which species are taking up those resources. Species-specific responses to spatial and temporal variation in non-resource environmental variables can also weaken interspecific interactions \citep{chesson_mechanisms_2000}, though in a previous study we failed to find evidence of a temporal storage effect in this community \citep{adler_weak_2009}. Interactions with above- or belowground enemies can regulate populations and, given enough species-specificity in the effects of these enemies, decouple interspecific dynamics \citep{hersh_evaluating_2011,janzen_herbivores_1970,connell_role_1971}.  More likely, plants may alter the environment, and even resource availability, in ways that facilitate their neighbors. For example, shade cast by \textit{Artemisia} canopies may reduce evapotranspiration (e.g. \citealt{Barbier2008}),

%adding these refs on hydraulic lift:

and \textit{Artemisia} may also increase soil moisture through hydraulic lift \citep{richards_hydraulic_1987}, and this may also increase nitrogen availability \citep{cardon_sagebrush_2013}. \textit{Artemisia} removal might offest this facilitative effect by reducing resource uptake, resulting in little net change in water balance following removal.  Unfortunately, we did not monitor soil moisture at the fine spatial and temporal resolution necessary to test this hypothesis.

% --------------------------------------------------------
% section on caldwell's resource competition studies: 
% --------------------------------------------------------

It is interesting to note, that the one grass species which showed a stronger than expected growth response to \textit{Artemisia} removal, \textit{P. spicata}, has been shown in a number of studies to compete strongly for soil resources with big sagebrush \textit{Artemisia tridentata}, a close relative of the shrub in our study \citep{caldwell_competition_1985,caldwell_rapid_1987,caldwell_exploitation_1991}. \textit{Artemisia tridentata} in these studies was shown to be more effective at competing with \textit{P. spicata} for phosphorous than with another perennial grass species \citep{caldwell_competition_1985}. Thus it seems plausible that reduced competition for soil nutrients may be responsible for the strong positive effect of \textit{Artemisia} removal on \textit{P. spicata} in our study.

@@

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% I added this nutnet tie-in that I think is cool; it could go up 
% after the Yenni reference above instead.   
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~`
